---
title: "ScientificProgrammingAssignment"
author: "Rita Sarantidou"
date: "9/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
DIR<-setwd("C:/Users/ritaa/Scientific Programming")
library(dplyr)
library(DMwR)
library(sm)
library(FactoMineR)
library(ggfortify)
library(viridis)  # for colouring
library(randomForest)
library(rfUtilities)
library(sp)
library(ROSE)
library(groupdata2)
library(ROCR)
library(UBL)
library(e1071) 
library(FSelector)
library(factoextra)
```

Importing the dataset, and transposing it to get 546 rows(individuals) and 47283 columns (probes).
In addition, import the individuals' info table (age,gender,condition, etc).

```{r}
d <- read.csv('Adjustedexpressionvalues.txt',row.names = 1, header = F)
d <-t(d)

# View(df) # observe the data frame
# dim(df)  # observe the dimensions of the dataset

individuals <- read.csv('datasets_451195_852422_E-MTAB-8018.sdrf.txt', sep = '')
individuals$part.<- as.character(individuals$part.)

ind1 <-individuals[,4:6]  # keep columns that contain individuals' ID, gender, and age
ind <- cbind(ind1, individuals[,10]) # combine above with individuals' condition (bipolar, normal, scizophrenic)
```


Merge the two data frames, i.e. gene expressions and individuals' information
```{r}
data <- merge(x = ind, y = d , by.x = "Characteristics.individual.", by.y = "NaN", all = F)
names(data)[names(data) == 'individuals[, 10]'] <- 'condition'

data <- data[order(data$condition),] # sorting the data frame based on the individuals' condition
sum(is.na(data))   # check how many missing values are in the data
table(data$condition)

data$condition <- as.character(data$condition)
data$condition[data$condition == "schizoaffective"]<-"schizophrenia"
#data[data$condition == "schizoaffective"]<-"schizophrenia"
#table(data$condition)

data$condition <- as.factor(data$condition)
table(data$condition)
```

Kennard and Stone method for partitioning

subsetting the data to bipolar (bd), normal (norm), schizoaffective (sca), and scizophrenic (scz)
```{r}
df <- data

bd <- subset(df, condition == "bipolar")
norm <- subset(df, condition == "normal")
scz <- subset(df, condition == "schizophrenia")
```

Density plots
```{r}
# for bipolar patients

x <- bd[,5:47287]
x<-knnImputation(x, k = 5, scale = T)
anyNA(x)

x <- as.numeric(unlist(x))

h<-hist(x, breaks=10, col="red", xlab="DEG values",
   main="Histogram with Normal Curve")
xfit<-seq(min(x),max(x),length=40)
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x))
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2)
```
```{r}
# for normal patients

n <- norm[,5:47287]
n<-knnImputation(n, k = 10, scale = T)
anyNA(n)

n <- as.numeric(unlist(n))

hn<-hist(n, breaks=10, col="red", xlab="DEG values",
   main="Histogram with Normal Curve")
nxfit<-seq(min(n),max(n),length=40)
nyfit<-dnorm(nxfit,mean=mean(n),sd=sd(n))
nyfit <- nyfit*diff(hn$mids[1:2])*length(n)
lines(nxfit, nyfit, col="blue", lwd=2)
```


```{r}
# for schizophrenic patients

s <- scz[,5:47287]
s<-knnImputation(s, k = 10, scale = T)
anyNA(s)

s <- as.numeric(unlist(s))

hs<-hist(s, breaks=10, col="red", xlab="DEG values",
   main="Histogram with Normal Curve")
sxfit<-seq(min(s),max(s),length=40)
syfit<-dnorm(sxfit,mean=mean(s),sd=sd(s))
syfit <- syfit*diff(hs$mids[1:2])*length(s)
lines(sxfit, syfit, col="blue", lwd=2)
```



Compare the 3 densities, run in console window
```{r}
# g<-factor(c("bipolar", "normal", "schizophrenic", "schizoaffective"))
# sm.density.compare(c(x,n,s,sa), g, band = F)
# colfill<-c(2:(2+length(levels(g))))
# legend(locator(1), levels(g), fill=colfill)
```

PCA analysis and visualization.

Since an eigenvalue <1 would mean that the component actually explains less than a single explanatory variable we would like to discard those.
If our data is well suited for PCA we should be able to discard these components while retaining at least 70–80% of cumulative variance.

```{r}
ddf <- knnImputation(df, k = 5, scale = T)  # imputing NAs 
anyNA(ddf)
ddf <- ddf[,5:47287]

ddf.pca <- prcomp(ddf, center = T, scale = T)
summary(ddf.pca)

eigen <- ddf.pca$sdev^2  # by squaring the eigenvalues we get the variance explained by each PC

perc_variance<- as.vector(eigen/sum(eigen)*100)

plot(cumsum(ddf.pca$sdev^2/sum(ddf.pca$sdev^2)), xlab = "PC #", ylab = "Amount of explained variance", main = "Cumulative variance plot") # plot cumulative explained variance
abline(v = 370, col="blue", lty=5)
abline(h = 0.80, col="blue", lty=5)
legend("topleft", legend=c("Cut-off @ PC370"),
       col=c("blue"), lty=5, cex=0.6)

# screeplot(ddf.pca, type = "l", npcs = 500, main = "Screeplot of the first 400 PCs")
# abline(h = 1, col="red", lty=5)
# legend("topright", legend=c("Eigenvalue = 1"),
#        col=c("red"), lty=5, cex=0.6)

```


PCA reduced data
```{r}
df_Reduced <- ddf.pca$x[,1:370]
df_Reduced <- cbind(df_Reduced, as.numeric(df$condition)-1)
colnames(df_Reduced)[371] <- "condition"
```

Partition the data in 80% training and 20% test sets.
Data splitting using the partition function
```{r}
set.seed(2)

df_Reduced<-as.data.frame(df_Reduced)

df_Reduced$condition<- as.character(df_Reduced$condition)

df_Reduced$condition[df_Reduced$condition == 0]<-"bipolar"
df_Reduced$condition[df_Reduced$condition == 1]<-"normal"
df_Reduced$condition[df_Reduced$condition == 2]<-"schizophrenia"
table(df_Reduced$condition)

df_Reduced$condition <- as.factor(df_Reduced$condition)

bd <- subset(df_Reduced, condition == "bipolar")
norm <- subset(df_Reduced, condition == "normal")
scz <- subset(df_Reduced, condition == "schizophrenia")

par <- rbind(bd,scz)

parts <- partition(par, p = 0.2, cat_col = 'condition') # split the bipolar and schizophrenic individuals

testset<-parts[[1]]
table(testset$condition)

trainset<-parts[[2]]
table(trainset$condition)


parts1 <- partition(norm, p = 0.2) # split the normal individuals, to randomly select a number of them for the test and training sets below
norm1 <- parts1[[1]]
norm2 <- parts1[[2]]

random_norm1 <-sample_n(norm1, 15)

random_norm2 <-sample_n(norm2, 200)

testset<-rbind(testset[1:16,],random_norm1, testset[17:30,])
table(testset$condition)
trainset <-rbind(trainset[1:68,],random_norm2,trainset[69:125,])
table(trainset$condition)

```

Balancing the classes using sythetic data for the training set.
ROSE (Random Over-Sampling Examples)
creates a sample of synthetic data by enlarging the features space of minority and majority class examples.

```{r}
set.seed(1)

balance_bipolar <- ROSE(condition~., data = trainset[1:268,], N = 400)
table(balance_bipolar$data$condition)
bp <-balance_bipolar$data
nm <- subset(bp, condition == "normal")
bpd <- subset(bp, condition == "bipolar")

balance_scz <- ROSE(condition~., data = trainset[69:325,], N = 400)
table(balance_scz$data$condition)
sc <-balance_scz$data

nm2 <- subset(sc, condition == "normal")
sch <- subset(sc, condition == "schizophrenia")

nm3 <- sample_n(nm, 100)  # randomly select 100 rows
nm4 <- sample_n(nm2,100)

NM <- rbind(nm3, nm4)  # bind the normal subsets

trainset_ROSE <- rbind(bpd,NM, sch)  # Bind the bipolar, normal, and schizophrenic subsets
table(trainset_ROSE$condition)

```

ADASYN for balancing the classes samples and splitting the data in 80% train 20% test sets
```{r}
 set.seed(7)
 ada <- as.data.frame(trainset)
 ada$condition <- as.factor(ada$condition)
 table(ada$condition)
 
 trainset_ADA <- AdasynClassif(condition~., ada, beta = 1, baseClass = "normal")
 table(trainset_ADA$condition)
 
# trainset_ADA$condition <- as.character(trainset_ADA$condition)
# 
# trainset_ADA$condition[trainset_ADA$condition == 0]<-"bipolar"
# trainset_ADA$condition[trainset_ADA$condition == 1]<-"normal"
# trainset_ADA$condition[trainset_ADA$condition == 2]<-"schizophrenia"
# 
# trainset_ADA$condition <- as.factor(trainset_ADA$condition)
# table(trainset_ADA$condition)
```


undersampling the normal (majority) class
```{r}
set.seed(123)
nm <- subset(trainset, condition == "normal")
bpd <- subset(trainset, condition == "bipolar")
sc <- subset(trainset, condition == "schizophrenia")

under_subset <-rbind(bpd,nm)

trainset_un <- ovun.sample(condition~., under_subset, method="under", N = 130)
trainset_under <- rbind(trainset_un$data,sc)
table(trainset_under$condition)

```



Random Forest classification for ROSE created data

```{r}
# Using For loop to identify the right mtry for model
set.seed(3)
a=c()
i=5
for (i in 3:20) {
  model <- randomForest(condition ~ ., data = trainset_ROSE, ntree = 500, mtry = i, importance = TRUE)
  predValid <- predict(model, testset_ROSE, type = "class")
  a[i-2] = mean(predValid == testset_ROSE$condition)
}
a

plot(3:20 ,a)
```

```{r}
#Tuning mtry parameter 

set.seed(4)
bestmtry <- tuneRF(trainset_ROSE, trainset_ROSE$condition, mtryStart = 3, stepFactor=1.5, improve=1e-5, ntree=500)
print(bestmtry)
```


nodesize 5:10
```{r}
set.seed(5)

node<-c(1:10)
acc<-c()

for (i in 1:10){
   
rf <- randomForest(condition ~ ., data=trainset_ROSE, ntree = 500, mtry= 19, importance = T, nodesize = node[i])
#rf <- rfcv(trainset, trainset$condition, cv.fold = 5)
prediction_for_rose <- predict(rf,testset[,-371])
acc[i] <-mean(prediction_for_rose == testset$condition)
}
#acc

best_node <- match(max(acc),acc) # to get the best node size

```

Building RF on ROSE training data
```{r}
set.seed(55)

rf <- randomForest(condition ~ ., data=trainset_ROSE, ntree = 500, mtry= 19, importance = T, nodesize = best_node)

rf

varImpPlot(rf, n.var = 5)

prediction_for_rose <- predict(rf,testset[,-371])

ac <-mean(prediction_for_rose == testset$condition)
```

# Here we train the random forest using the undersampled training set
```{r}
set.seed(555)
rf_under <- randomForest(condition ~ ., data=trainset_under, ntree = 500, mtry= 19, importance = T, nodesize = best_node)
prediction_for_under <- predict(rf_under,testset[,-371])

varImpPlot(rf_under, n.var = 5)
ac_under <-mean(prediction_for_under == testset$condition)

ac
ac_under

```

ROC curves and AUC
```{r}
set.seed(6)
# Calculate the probability of new observations belonging to each class
# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
prediction_for_roc_curve <- predict(rf,testset[,-371],type="prob")
# Use pretty colours:
pretty_colours <- c("#F8766D","#00BA38","#619CFF")
# Specify the different classes 
classes <- levels(testset$condition)
# For each class
for (i in 1:3)
{
 # Define which observations belong to class[i]
 true_values <- ifelse(testset[,371]==classes[i],1,0)
 # Assess the performance of classifier for class[i]
 pred <- prediction(prediction_for_roc_curve[,i],true_values)
 perf <- performance(pred, "tpr", "fpr")
 if (i==1)
 {
     plot(perf,main="ROC Curve RF ROSE",col=pretty_colours[i]) 
 }
 else
 {
     plot(perf,main="ROC Curve RF ROSE",col=pretty_colours[i],add=TRUE) 
 }

 # Calculate the AUC and print it to screen
 auc.perf <- performance(pred, measure = "auc")
 print(auc.perf@y.values)
}
 legend(0.6, 0.4, legend=classes,
       col=pretty_colours, lty=1:2, cex=0.8)
```


```{r}
set.seed(66)
# Calculate the probability of new observations belonging to each class
# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
prediction_for_roc_curve <- predict(rf_under,testset[,-371],type="prob")
# Use pretty colours:
pretty_colours <- c("#F8766D","#00BA38","#619CFF")
# Specify the different classes 
classes <- levels(testset$condition)
# For each class
for (i in 1:3)
{
 # Define which observations belong to class[i]
 true_values <- ifelse(testset[,371]==classes[i],1,0)
 # Assess the performance of classifier for class[i]
 pred <- prediction(prediction_for_roc_curve[,i],true_values)
 perf <- performance(pred, "tpr", "fpr")
 if (i==1)
 {
     plot(perf,main="ROC Curve RF undersampled",col=pretty_colours[i]) 
 }
 else
 {
     plot(perf,main="ROC Curve RF undersampled",col=pretty_colours[i],add=TRUE) 
 }

 # Calculate the AUC and print it to screen
 auc.perf <- performance(pred, measure = "auc")
 print(auc.perf@y.values)
}
 legend(0.6, 0.4, legend=classes,
       col=pretty_colours, lty=1:2, cex=0.8)
```


Random Forest on ADASYN data

```{r}
set.seed(8)
#Tuning mtry parameter 

bestmtry_ada <- tuneRF(trainset_ADA, trainset_ADA$condition, mtryStart = 3, stepFactor=1.5, improve=1e-5, ntree=500)
print(bestmtry_ada)
```

```{r}
set.seed(9)

# node<-c(1:10)
ACC<-c()

for (i in 1:10){
   
rf <- randomForest(condition ~ ., data=trainset_ADA, ntree = 500, mtry= 19, importance = T, nodesize = node[i])
#rf <- rfcv(trainset, trainset$condition, cv.fold = 5)
prediction_for_ada <- predict(rf,testset[,-371])
ACC[i] <-mean(prediction_for_ada == testset$condition)
}
#acc

best_n <- match(max(ACC),ACC) # to get the best node size



rf_ada <- randomForest(condition ~ ., data=trainset_ADA, ntree = 500, mtry= 28, importance = T, nodesize=best_n)
#rf <- rfcv(trainset, trainset$condition, cv.fold = 5)
rf_ada

varImpPlot(rf_ada, n.var = 5)

prediction_for_ada <- predict(rf_ada,testset[,-371])
#table(observed=testset_ADA$condition,predicted=prediction_for_ada)

#  
# testset$condition <- as.character(testset$condition)
# 
# testset$condition[testset$condition == 0]<-"bipolar"
# testset$condition[testset$condition == 1]<-"normal"
# testset$condition[testset$condition == 2]<-"schizophrenia"


Accuracy_ada <- mean(prediction_for_ada == testset$condition)
Accuracy_ada
# Accuracy_rose <- mean(prediction_for_rose== testset_ROSE$condition)
# Accuracy_rose


```

ROC curves and AUC for ADASYN data
```{r}
set.seed(10)
# Calculate the probability of new observations belonging to each class
# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
prediction_for_roc_ada <- predict(rf_ada,testset[,-371],type="prob")
# Use pretty colours:
pretty_colours <- c("#F8766D","#00BA38","#619CFF")

# Specify the different classes 
classes <- levels(testset$condition)

# For each class
for (i in 1:3)
{
 # Define which observations belong to class[i]
 true_values <- ifelse(testset[,371]==classes[i],1,0)
 # Assess the performance of classifier for class[i]
 pred <- prediction(prediction_for_roc_ada[,i],true_values)
 perf <- performance(pred, "tpr", "fpr")
 if (i==1)
 {
     plot(perf,main="ROC Curve RF ADASYN",col=pretty_colours[i]) 
 }
 else
 {
     plot(perf,main="ROC Curve RF ADASYN",col=pretty_colours[i],add=TRUE) 
 }

 # Calculate the AUC and print it to screen
 auc.perf <- performance(pred, measure = "auc")
 print(auc.perf@y.values)
}
 legend(0.6, 0.4, legend=classes,
       col=pretty_colours, lty=1:2, cex=0.8)

```



SVM tuning parameters
```{r}
# Tune SVM cost and gamma parameters for ROSE generated samples
set.seed(11)
tune_ROSE <- tune.svm(condition~., data = trainset_ROSE, type="C-classification", kernel="radial", gamma = c(0.001, 0.01, 0.1), cost = c(0.5, 1, 5, 10, 15) ) 

tune_ROSE$best.parameters$gamma
tune_ROSE$best.parameters$cost
```

```{r}
# Tune SVM cost and gamma parameters for ADASYN generated samples
set.seed(12)
tune_ada <- tune.svm(condition~., data = trainset_ADA, type="C-classification", kernel="radial", gamma = c(0.001, 0.01, 0.1), cost = c(0.5, 1, 5, 10, 15) ) 

tune_ada$best.parameters$gamma
tune_ada$best.parameters$cost
```

```{r}
# Tune SVM cost and gamma parameters for undersampling
set.seed(122)
tune_under <- tune.svm(condition~., data = trainset_under, type="C-classification", kernel="radial", gamma = c(0.001, 0.01, 0.1), cost = c(0.5, 1, 5, 10, 15) ) 

tune_under$best.parameters$gamma
tune_under$best.parameters$cost
```

Building SVM models
```{r}
set.seed(88)

svm_ROSE <- svm(condition~.,
                data = trainset_ROSE,
               type = "C-classification", kernel = "radial", gamma = tune_ROSE$best.parameters$gamma,
               cost = tune_ROSE$best.parameters$cost,
               probability = T,
               cross = 5
             )

pred_svm_ROSE <- predict(svm_ROSE, testset)
Accuracy_svm_ROSE <- mean(pred_svm_ROSE == testset$condition)
Accuracy_svm_ROSE
summary(svm_ROSE)
```

```{r}
set.seed(99)

svm_ada <- svm(condition~.,
                data = trainset_ADA,
               type = "C-classification", kernel = "radial", gamma = tune_ada$best.parameters$gamma,
               cost = tune_ada$best.parameters$cost,
               probability = T,
               cross = 5
             )

pred_svm_ada <- predict(svm_ada, testset)
Accuracy_svm_ada <- mean(pred_svm_ada == testset$condition)
Accuracy_svm_ada
summary(svm_ada)
```

```{r}
set.seed(100)

svm_under <- svm(condition~.,
                data = trainset_under,
               type = "C-classification", kernel = "radial", gamma = tune_under$best.parameters$gamma,
               cost = tune_under$best.parameters$cost,
               probability = T,
               cross = 5
             )

pred_svm_under <- predict(svm_under, testset)
Accuracy_svm_under <- mean(pred_svm_under == testset$condition)
Accuracy_svm_under
summary(svm_under)
```

Plot ROC for SVM
```{r}
set.seed(13)
# Calculate the probability of new observations belonging to each class
# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
prediction_for_roc_svm <- predict(svm_ROSE,testset,type="prob", probability = T)
svm_attr <- attr(prediction_for_roc_svm, "probabilities")

# Use pretty colours:
pretty_colours <- c("#F8766D","#00BA38","#619CFF")

# Specify the different classes 
classes <- levels(testset$condition)

# For each class
for (i in 1:3)
{
 # Define which observations belong to class[i]
 true_values <- ifelse(testset[,371]==classes[i],1,0)
 # Assess the performance of classifier for class[i]
 
 pred <- prediction(svm_attr[,i],true_values)
 perf <- performance(pred, "tpr", "fpr")
 if (i==1)
 {
     plot(perf,main="ROC Curve SVM ROSE",col=pretty_colours[i]) 
 }
 else
 {
     plot(perf,main="ROC Curve SVM ROSE",col=pretty_colours[i],add=TRUE) 
 }

 # Calculate the AUC and print it to screen
 auc.perf <- performance(pred, measure = "auc")
 print(auc.perf@y.values)
}

 legend(0.6, 0.4, legend=classes,
       col=pretty_colours, lty=1:2, cex=0.8)

```


```{r}
set.seed(14)
# Calculate the probability of new observations belonging to each class
# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
prediction_for_roc_svm <- predict(svm_ada,testset,type="prob", probability = T)
svm_attr <- attr(prediction_for_roc_svm, "probabilities")

# Use pretty colours:
pretty_colours <- c("#F8766D","#00BA38","#619CFF")

# Specify the different classes 
classes <- levels(testset$condition)

# For each class
for (i in 1:3)
{
 # Define which observations belong to class[i]
 true_values <- ifelse(testset[,371]==classes[i],1,0)
 # Assess the performance of classifier for class[i]
 
 pred <- prediction(svm_attr[,i],true_values)
 perf <- performance(pred, "tpr", "fpr")
 if (i==1)
 {
     plot(perf,main="ROC Curve SVM ADASYN",col=pretty_colours[i]) 
 }
 else
 {
     plot(perf,main="ROC Curve SVM ADASYN",col=pretty_colours[i],add=TRUE) 
 }

 # Calculate the AUC and print it to screen
 auc.perf <- performance(pred, measure = "auc")
 print(auc.perf@y.values)
}

 legend(0.6, 0.4, legend=classes,
       col=pretty_colours, lty=1:2, cex=0.8)

```

```{r}
set.seed(144)
# Calculate the probability of new observations belonging to each class
# prediction_for_roc_curve will be a matrix with dimensions data_set_size x number_of_classes
prediction_for_roc_svm <- predict(svm_under,testset,type="prob", probability = T)
svm_attr <- attr(prediction_for_roc_svm, "probabilities")

# Use pretty colours:
pretty_colours <- c("#F8766D","#00BA38","#619CFF")

# Specify the different classes 
classes <- levels(testset$condition)

# For each class
for (i in 1:3)
{
 # Define which observations belong to class[i]
 true_values <- ifelse(testset[,371]==classes[i],1,0)
 # Assess the performance of classifier for class[i]
 
 pred <- prediction(svm_attr[,i],true_values)
 perf <- performance(pred, "tpr", "fpr")
 if (i==1)
 {
     plot(perf,main="ROC Curve SVM undersampling",col=pretty_colours[i]) 
 }
 else
 {
     plot(perf,main="ROC Curve SVM undersampling",col=pretty_colours[i],add=TRUE) 
 }

 # Calculate the AUC and print it to screen
 auc.perf <- performance(pred, measure = "auc")
 print(auc.perf@y.values)
}

 legend(0.6, 0.4, legend=classes,
       col=pretty_colours, lty=1:2, cex=0.8)

```

Visualizing PCA

```{r}

fviz_eig(ddf.pca)  # Visualize eigenvalues (scree plot). Show the percentage of variances explained by each principal component.

groups <- df$condition
fviz_pca_ind(ddf.pca,
             col.ind = groups, # color by groups
             palette = viridis(4),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups",
             repel = TRUE
             )
```

```{r}
fviz_pca_var(ddf.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```

```{r}
set.seed(123)

PCAcolors <- c(viridis(3))[as.integer(groups)]
PCAscores <- ddf.pca$x
PCAloadings <- ddf.pca$rotation

par(mfrow=c(1,1))
plot(PCAscores[,1:2],  # x and y data
     pch=4,           # point shape
     col=PCAcolors,    # point border color
     bg=PCAcolors,     # point color
     cex=1,          # point size
     main="Scores"     # title of plot
)

legend("topright",                                # position of legend
       legend=levels(groups),                       # legend display
       pch=4,                                    # point shape
       pt.bg=c("#440154FF", "#21908CFF", "#FDE725FF"),    # point colors
       pt.cex=1.5,                                # point size
       col = c("#440154FF", "#21908CFF", "#FDE725FF")    # point border color
)

plot(PCAloadings[,1:2],   # x and y data
     pch=21,              # point shape
     bg="black",          # point color
     cex=0.1,               # point size
     main="Loadings"      # title of plot
)
text(PCAloadings[,1:2],             # sets position of labels
     labels=rownames(PCAloadings)   # print labels
)
```

Plot important PCs based on the randomf forest classification
```{r}

loadings <- ddf.pca$rotation

lim1 <- subset(loadings[,38], loadings[,38]> 0.02)
lim2 <- subset(loadings[,38], loadings[,38]< -0.02)
labels_PC38 <- c(as.vector(names(lim1)), as.vector(names(lim2)))

plot(ddf.pca$rotation[,38], 
     main = "PC38",
     ylab = "Variable Loadings",
     xlab = "DEG")
abline(h = 0.02, col = "red")
abline(h = -0.02, col = "red")
# text(x= 0:length(ddf.pca$rotation[,38]), 
#      labels = names(lim1), 
#      pos =3, cex = 0.2, col = 'blue')

# For PC43
lim3 <- subset(loadings[,43], loadings[,43]> 0.02)
lim4 <- subset(loadings[,43], loadings[,43]< -0.02)
labels_PC43 <- c(as.vector(names(lim3)), as.vector(names(lim4)))

plot(ddf.pca$rotation[,43], 
     main = "PC43",
     ylab = "Variable Loadings",
     xlab = "DEG")
abline(h = 0.02, col = "red")
abline(h = -0.02, col = "red")
# text(x= 0:length(ddf.pca$rotation[,43]), 
#      labels = names(lim1), 
#      pos =3, cex = 0.2, col = 'blue')

# Get the important DEG labels/IDs that are in both PCs
imp_IDs <- intersect(labels_PC38,labels_PC43)
imp_IDs
```



